\subsection{Logging a UserTag event}
\label{log_usertag}

User tag is an arbitrary ``name=value'' pair with which the user can assign
additional information to a job. Further on, LB can be queried based also on
values of user tags.  % (see Sect.~\ref{tag-query}).

In order to add user tag for a job a special event \verb'UserTag' is used. This
event can be logged by the job owner using the glite-lb-logevent command (see
also sec.\ref{glite-lb-logevent}). Here we suppose the command is used from
user's running application because a correct setting of environment variables
needed by the command is assured.

General template for adding user tag is as follows:

\begin{verbatim}
glite-lb-logevent -s Application -e UserTag    \
        -j <job_id>                         \
        -c <seq_code>                       \
        --name <tag_name>                   \
        --value <tag_value>
\end{verbatim}

where

\begin{tabularx}{\textwidth}{lX}
\verb'<job_id>'    & specifies the job to change \\
\verb'<seq_code>'  & specifies the sequence code returned by previous call
			of verb 'glite-lb-logevent' \\
\verb'<tag_name>'  & specifies the name of user tag \\
\verb'<tag_value>' & specifies the value of user tag \\
\end{tabularx}

The user application is always executed from within a JobWrapper script (part
of Workload Management System \cite{jgc}). The wrapper  sets the  appropriate
\verb'JobId' in the environment variable \verb'GLITE_WMS_JOBID'. The user
should pass this value to the \verb'-j' option of \verb'glite-lb-logevent'.
Similarly, the wrapper sets an initial value of the event sequence code in the
environment variable \verb'GLITE_WMS_SEQUENCE_CODE'.

If the user application calls \verb'glite-lb-logevent' just once, it is
sufficient to pass this value to the \verb'-c' option.  However, if there are
more  subsequent calls,  the  user is responsible for capturing an updated
sequence code from the stdout of \verb'glite-lb-logevent' and using it in
subsequent calls. The \LB\ design requires the sequence codes in  order  to be
able to sort events correctly while not relying on strictly synchronized
clocks.  

\TODO{What are the possible values/types of the tag? String/Integer/Float/Double?}

The example bellow is a job consisting of 100 phases. A user tag phase is used
to log the phase currently being executed. Subsequently, the user may monitor
execution of the job phases as a part of the job status returned by \LB.

\begin{verbatim}
  #!/bin/sh

  for p in `seq 1 100`; do

  # log the UserTag event
  GLITE_WMS_SEQUENCE_CODE=`glite-lb-logevent -s Application
    -e UserTag
    -j $GLITE_WMS_JOBID -c $GLITE_WMS_SEQUENCE_CODE
    --name=phase --value=$p`

  # do the actual computation here
  done
\end{verbatim}

\TODO{Add more examples.}

