%
%% Copyright (c) Members of the EGEE Collaboration. 2004-2010.
%% See http://www.eu-egee.org/partners for details on the copyright holders.
%% 
%% Licensed under the Apache License, Version 2.0 (the "License");
%% you may not use this file except in compliance with the License.
%% You may obtain a copy of the License at
%% 
%%     http://www.apache.org/licenses/LICENSE-2.0
%% 
%% Unless required by applicable law or agreed to in writing, software
%% distributed under the License is distributed on an "AS IS" BASIS,
%% WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
%% See the License for the specific language governing permissions and
%% limitations under the License.
%
\subsection{glite-lb-notify}
\label{s:lb-notify}

\verb'glite-lb-notify' is a fairly simple wrapper on the \LB notification API
(see also \cite{lbdg}).
It allows to create a notification (with a restricted richness of specifying 
conditions), bind to it for receiving notifications, and drop it finally.

\LB notification is a user-initiated trigger at the server.
Whenever a job enters a state matching conditions specified with the notification,
the current state of the job is sent to the notification client.
On the other hand, the notification client is a network listener
which receives server-initiated connections to get the notifications.
Unless \verb'-s' is specified, the notification library creates the listener
socket.

Within the notification validity, clients can disappear and even migrate.
However, only a single active client for a notification is allowed.

\LB server and port to contact must be specified with GLITE\_WMS\_NOTIF\_SERVER 
environment variable.

\verb'glite-lb-notify' is supported by \LBver{2.x} and newer. In \LBver{1.x}, \verb'glite-lb-notify' 
with very limited functionality can be found in \verb'examples' directory.

\verb'glite-lb-notify' support these actions:

\begin{tabularx}{\textwidth}{lX}
\texttt{new} & Create new notification registration.\\
\texttt{bind} &  Binds an notification registration to a client.\\
\texttt{refresh} &  Enlarge notification registration validity.\\
\texttt{receive}  & Binds to an existing notification registration and listen to
server.\\
\texttt{drop}     & Drop the notification registration.\\
\end{tabularx}

For action \verb'new', command usage is:

\begin{verbatim}
  glite-lb-notify new [ { -s socket_fd | -a fake_addr } -t requested_validity
           -j jobid  { -o owner | -O }  -n network_server 
           -v virtual_organization --states state1,state2,... -c -f flags]
\end{verbatim}

For action \verb'bind', command usage is:
\begin{verbatim}
  glite-lb-notify bind [ { -s socket_fd | -a fake_addr } -t requested_validity ] 
           notifid
\end{verbatim}

For action \verb'refresh', command usage is:
\begin{verbatim}
  glite-lb-notify refresh [-t requested_validity ] notifid
\end{verbatim}

For action \verb'receive', command usage is:
\begin{verbatim}
  glite-lb-notify receive  [ { -s socket_fd | -a fake_addr } ] [-t requested_validity ] [-i timeout] [-r ] [-f field1,field2,...] [notifid]
\end{verbatim}

For action \verb'drop', command usage is:
\begin{verbatim}
   glite-lb-notify drop notifid
\end{verbatim}

where

\begin{tabularx}{\textwidth}{lX}
\texttt{  notifid} & Notification ID \\
\texttt{  -s socket\_fd} &  allows  to  pass  a opened listening socket  \\
\texttt{  -a fake\_addr} &  fake the client address \\
\texttt{  -t requested\_validity} & requested validity of the notification (in seconds)   \\
\texttt{  -j jobid} & job ID to connect notification registration with   \\
\texttt{  -o owner} & match this owner DN   \\
\texttt{  -O} & match owner on current user's DN \\
\texttt{  -n network\_server} &  match only this network server (WMS entry point)  \\
\texttt{  -v virtual\_organization} & match only jobs of this Virtual Organization  \\
\texttt{  -i timeout} & timeout to receive operation in seconds   \\
\texttt{  -f field1,field2,...} & list of status fields to print (only owner by default)   \\
\texttt{  -c} & notify only on job state change \\
\texttt{  -S, --state state1,state2,...} & match on events resulting in listed states   \\
\texttt{  -r} & refresh automatically the notification registration while receiving data\\
\end{tabularx}

For additional information see also manual page glite-lb-notify(1).

\subsubsection{Example: Registration and waiting for simple notification}
\label{e:notify}

Following steps describe basic testing procedure of the notification
system by registering a notification on any state change of this job
and waiting for notification.

Register notification for a given jobid
%(\verb'https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q'), 
with validity 2 hours (7200 seconds):

\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify new \
   -j https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q -t 7200
\end{verbatim}

returns:

\begin{verbatim}
  notification ID: https://skurut68-2.cesnet.cz:9100/NOTIF:tOsgB19Wz-M884anZufyUw 
\end{verbatim}


Wait for notification (with timeout 120 seconds):
\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify receive \
   -i 120 https://skurut68-2.cesnet.cz:9100/NOTIF:tOsgB19Wz-M884anZufyUw 
\end{verbatim}

returns:
\begin{verbatim}
  notification is valid until: '2008-07-29 15:04:41' (1217343881)
  https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q        Waiting
      /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
  https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q        Ready
      /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
  https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q        Scheduled
      /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
  https://skurut68-2.cesnet.cz:9100/D1qbFGwvXLnd927JOcja1Q        Running
      /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
\end{verbatim}

Destroy notification:
\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify drop \
    https://skurut68-2.cesnet.cz:9100/NOTIF:tOsgB19Wz-M884anZufyUw
\end{verbatim}


\subsubsection{Example: Waiting for notifications on all user's jobs}

Instead of waiting for one job, user may want to accept notification about 
state changes of all his jobs.

Register notification for actual user:
\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify new -O
\end{verbatim}

returns:

\begin{verbatim}
  notification ID: https://skurut68-2.cesnet.cz:9100/NOTIF:tOsgB19Wz-M884anZufyUw 
\end{verbatim}

And continue with \verb'glite-lb-notify receive' similarly to previous example.
But in this case, we want to display also other information about job --
not job owner, but destination (where job is running) and location (which component is 
serving job):

\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify receive \
   -i 240 -f destination,location \
   https://skurut68-2.cesnet.cz:9100/NOTIF:tOsgB19Wz-M884anZufyUw
\end{verbatim}

returns:

\begin{verbatim}
  notification is valid until: '2008-07-29 15:43:41' (1217346221)

 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   (null) NetworkServer/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   (null)  destination queue/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   (null) WorkloadManager/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   (null)  name of the called component/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   destination CE/queue WorkloadManager/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Waiting
   destination CE/queue WorkloadManager/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Ready
   destination CE/queue destination queue/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Ready
   destination CE/queue JobController/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Ready
   destination CE/queue LRMS/destination hostname/destination instance
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Ready
   destination CE/queue LogMonitor/erebor.ics.muni.cz/
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Scheduled
   destination CE/queue LRMS/destination hostname/destination instance
 https://skurut68-2.cesnet.cz:9100/qbRFxDFCg2qO4-9WYBiiig        Running
   destination CE/queue LRMS/worknode/worker node

\end{verbatim}

\subsubsection{Example: Registering for notifications to be delivered over ActiveMQ}
\label{e:notifymsg}

Delivering notification messages over the messaging infrastructure provided by ActiveMQ is a feature introduced in \LBver{3.0}. It uses the fake address capability (\texttt{-a} option) to specify messaging topic to use when generating messages.

\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify new \
   -O -a x-msg://grid.emi.lbexample
\end{verbatim}

Rather than using the \LB notification API to receive messages, tap to the given messaging topic (\texttt{grid.emi.lbexample} in our case) in the messaging infrastructure.

Note that production environments can impose restrictions on topic names. In the context of EGI, for instance, the ``\texttt{grid.emi.}'' prefix is mandatory.

\subsubsection{Example: Waiting for more notifications on one socket}

The foloving example demonstrates possibility to reuse existing socket for receiving
multiple notifications. Perl script \verb'notify.pl' (available in 
examples directory) creates socket, which is then reused for all
\verb'glite-lb-notify' commands.

\begin{verbatim}
GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 NOTIFY_CMD=glite-lb-notify \
 ./notify.pl -O
\end{verbatim}

returns:

\begin{verbatim}
notification ID: https://skurut68-2.cesnet.cz:9100/NOTIF:EO73rjsmexEZJXuSoSZVDg
valid: '2008-07-29 15:14:06' (1217344446)
got connection
https://skurut68-2.cesnet.cz:9100/ANceuj5fXdtaCCkfnhBIXQ        Submitted
/DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
glite-lb-notify: Connection timed out (read message)
got connection
https://skurut68-2.cesnet.cz:9100/p2jBsy5WkFItY284lW2o8A        Submitted
/DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
glite-lb-notify: Connection timed out (read message)
got connection
https://skurut68-2.cesnet.cz:9100/p2jBsy5WkFItY284lW2o8A        Waiting
/DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
\end{verbatim}


\subsubsection{Example: Waiting for notifications on jobs reaching terminal states}

This example shows how to set up notifications for jobs reaching state \emph{done} or \emph {aborted}. Since we prefer to receive just one notification per job, we will also use the \texttt{-c} option, which makes sure that notifications will be generated only on actual job state change.


\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify new \
   --state done,aborted -c
\end{verbatim}

returns:

\begin{verbatim}
  notification ID: https://skurut68-2.cesnet.cz:9100/NOTIF:6krjMRshTouH2n7D9t-xdg 
  valid: '2009-04-30 06:59:18 UTC' (1241074758)	
\end{verbatim}

Wait for notification (with timeout 120 seconds):
\begin{verbatim}
  GLITE_WMS_NOTIF_SERVER=skurut68-2.cesnet.cz:9100 glite-lb-notify receive \
   -i 120 https://skurut68-2.cesnet.cz:9100/NOTIF:6krjMRshTouH2n7D9t-xdg 
\end{verbatim}

returns:
\begin{verbatim}
https://skurut68-2.cesnet.cz:9100/eIbQNz3oHpv-OkYVu-cXNg	Done
    /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
https://skurut68-2.cesnet.cz:9100/GpBy2gfIZOAXR2hKOAYGgg	Aborted
    /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
https://skurut68-2.cesnet.cz:9100/KWXmsqvsTQKizQ4OMiXItA	Done
    /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
https://skurut68-2.cesnet.cz:9100/O1zs50Nm1r03vx2GLyaxQw	Done
    /DC=cz/DC=cesnet-ca/O=Masaryk University/CN=Miroslav Ruda
\end{verbatim}

